The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}

### 2017-08-16 GRB111117A.tex
Addressing referee part 1

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed
at 21\,000~\AA~where the night sky is very bright. In addition, several bright
sky-lines are superposed on the line, making an accurate estimate of the
\ha-flux difficult. We obtain a limit on the SFR by numerically integrating the
part of \ha~free of contamination and correcting for the missing flux using the
line profile and obtain $F_{H \alpha} > 4.1 \times 10^{-17} \mathrm{erg}
\mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting the \citet{Kennicutt1998}
relation to a \citet{Chabrier2003} initial mass function using
\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$.
Additionally, based on the width of \oiii, an aperture covering the line is
integrated over, where synthetic sky lines \citep{Noll2012, Jones2013} have been
masked and interpolated over. From the integrated \ha-line, we estimate SFR =
$18 \pm 3$ M$_\odot$ yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and the
detection of \lya, the host is constrained to contain very little or no dust,
although the presence of Lya does not exclude dust. Therefore we do not apply a
dust-correction to the measured \ha~flux here. \oii~is close to a region of
strong telluric absorption, which is why no SFR is inferred from this line.

@@

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
faded. These new data are complemented by a re-analysis of some of the imaging
used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, see
Sect. \ref{xray}, there is likely no afterglow contamination on the measured
photometry.

All data were reduced, analyzed and fitted in a similar manner as described in
detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
reduction which includes bias/flat-field correction, de-fringing (if necessary),
sky-subtraction, and stacking of individual images. The photometry of the host
was calibrated relative to field stars from the SDSS and 2MASS catalogs in the
case of $grizJHK_{\mathrm{s}}$ filters.
For the $R$ and $I$-band photometry, we used the color transformations of
Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
We convert all magnitudes into the AB system, and correct for a Galactic 
foreground of $E_{B-V}=0.027~\mathrm{mag}$ \citep{Schlegel1998, Schlafly2011}.

@@

The multi-color SED is fit using the \citet{Bruzual2003} single stellar population
models based on a \citet{Chabrier2003} with initial mass function in
\emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to the
spectroscopic value of $z=2.211$. The best fit model is an unreddened galaxy
template, and returns physical parameters of absolute magnitude
($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) = 9.9\pm0.2$),
stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and star-formation rate
($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). We show the SED
fit in Fig.~\ref{fig:SED}.

Noteworthy is the discrepancy of our new VLT/FORS2 photometry and the
re-analysis of the Gemini data to the $z$-band measurements of
\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
$z$-band photometry that is brighter by 0.8~mag to 1.0 mag compared to our
value, where data taken in bluer filters are in excellent agreement. The large
$i-z$ color was mistakenly interpreted as a 4000\,\AA\,break driving the galaxy
photometric redshift of the earlier works. Using the revised photometry from
Table~\ref{tab:phot_overview}, the photometric redshift of the galaxy is
$z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the spectroscopic
value at the 1~$\sigma$ confidence level.

>>>>

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at
21\,000~\AA~where the night sky is very bright. In addition, several bright
sky-lines are superposed on the line, making an accurate estimate of the
\ha-flux difficult. Due to their velocity structure, the lines exhibit clear
deviations from a Gaussian. Given the low S/N of the spectra we do not attempt
any parametric fits. We thus obtain a limit on the SFR by numerically
integrating the part of \ha~free of contamination and obtain $F_{H \alpha} > 4.1
\times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting
the \citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass
function using \citet{Madau2014}, we derive a limit of $SFR > 7
M_\odot$~yr$^{-1}$. 
%Additionally, based on the width of \oiii, an aperture
%covering \ha is integrated over, where sky lines have been masked and
%interpolated over, using a synthetic sky spectrum \citep{Noll2012, Jones2013}.
%From the integrated \ha-line, we estimate SFR = $18 \pm 3$ M$_\odot$ yr$^{-1}$.
%The error on the SFR is derived from the integrated variance spectrum over the
%\ha-line and excluding the intrinsic scatter of the relation. 
From the SED-fit
(Sect.~\ref{SED}), and the detection of \lya, the host is constrained to contain
very little or no dust, although the presence of Lya does not exclude dust.
Therefore we do not apply a dust-correction to the measured \ha~flux here.
\oii~is close to a region of strong telluric absorption, which is why no SFR is
inferred from this line.

@@

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
faded. These new data are complemented by a re-analysis of some of the imaging
used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}. Most data were taken long after the GRB when no
afterglow contribution was present. Given the faintness of the afterglow (see
Sect. \ref{xray}, and \citealt{Cucchiara2011, Cenko2011}), we also expect
negligible contribution to the earliest epochs, which is confirmed by the
consistency between the two $g$-band measurements.

All data were reduced, analyzed and fitted in a similar manner as described in
detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
reduction which includes bias/flat-field correction, de-fringing (where
necessary), sky-subtraction, and stacking of individual images. On the final
reduced products, seeing-matched aperture photometry was carried out using
\texttt{Source Extractor} \citep{Bertin1996} where the size of the aperture was
chosen to be twice the seeing FWHM. (THOMAS! WAS THE APERTURE FIXED BETWEEN
FILTERS AND IS THIS CORRECT?) Photometric calibration was fixed relative to
field stars from the SDSS and 2MASS catalogs in the case of $griz$ and
$JHK_{\mathrm{s}}$ filters, respectively. For the $R$- and $I$-band photometry,
we used the color transformations of Lupton%
\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. We
convert all magnitudes into the AB system, and correct for a Galactic foreground
of $E_{B-V}=0.027~\mathrm{mag}$ \citep{Schlegel1998, Schlafly2011}.

Noteworthy is the discrepancy of both our new VLT/FORS2 photometry and the
re-analysis of the Gemini data compared to the $z$-band measurements of
\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
$z$-band magnitude that is brighter by 0.8-1.0 mag compared to our value, where
data taken in other filters are consistent within the errors. Visual inspection
of the Gemini image shows only a marginal detection of the host, which we report
here as a $\sim$ 2-$\sigma$ measurement. More conservatively, the 3-$\sigma$
upper limit for the Gemini image is $z > 24.06$. Objects of magnitude $z \sim
23.2$ are clearly seen and are significantly brighter than the GRB host galaxy.
The consistency between the deeper FORS2 $z$-band image and the upper limit
derived for the Gemini image lends credence to the inferred magnitudes, see
Fig. \ref{photplot}.

@@

The multi-color SED is fit using the \citet{Bruzual2003} single stellar
population models based on a \citet{Chabrier2003} with initial mass function in
\emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to the
spectroscopic value of $z=2.211$. For the SED fitting, we create a grid of
templates consisting of $\sim 10^6$  different galaxy templates with four
metallicities 0.02, 0.2, 0.4, 1.0 $Z_{\odot}$, different ages, star-formation
histories and degrees of extinction. For every model, we calculate the
likelihood, and create a probability density function (PDF) for a given
parameter by marginalizing over the other parameters. We quote the median of the
PDF as the best fit parameters and the errors are the 16th and 84th percentiles
of the PDFs \citep[e.g. see][for details on the SED fitting
procedure]{Schulze2016}.

The best fit model is an unreddened galaxy
template, and returns physical parameters of absolute magnitude
($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) = 9.9\pm0.2$),
stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and star-formation rate
($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). We show the SED
fit in Fig.~\ref{fig:SED}.

The large $i-z$ color was mistakenly interpreted as a 4000\,\AA~ break driving
the galaxy photometric redshift of the earlier works. Even without fixing the
redshift to the spectroscopic value, using the revised photometry from
Table~\ref{tab:phot_overview}, the photometric redshift of the galaxy is
$z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the spectroscopic
value at the 1~$\sigma$ confidence level.
